---
title: "Homework 3"
author: "Iris (Ziyi) Wang"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
#library(rnoaa)
library(ggridges)
library(patchwork)
#devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

knitr::opts_chunk$set(
    fig.width = 6, 
    fig.asp = .6,
    out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Due date

Due: October 10 at 10:00pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |
| Optional survey | No points |


### Problem 0

This "problem" focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files. 



### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)`  columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


### Problem 2

#### Load, tidy, and otherwise wrangle the accelerometer data. 
```{r accelerometer, message = F}
path_to_P2_data = "./data/accel_data.csv"
accelerometer = read_csv(path_to_P2_data) %>% 
    janitor::clean_names() %>% 
    pivot_longer(
        cols = starts_with("activity_"),
        names_to = 'activity_number',
        values_to = "activity_minute",
        names_prefix = "activity_") %>% 
    mutate(
        activity_minute = as.numeric(activity_minute),
        activity_number = as.numeric(activity_number),
        day = factor(day),
        # add weekday_vs_weekend variable
        week_end_vs_day = ifelse(day_id == c(3,4), "weekend", "weekday")
    )
skimr::skim_without_charts(accelerometer)
```

After tidy and wrangle the data, the resulting _**accelerometer**_ data contains `r nrow(accelerometer)` rows and `r ncol(accelerometer)` columns. The variables in the _**accelerometer**_ are: `r names(accelerometer)`. Specifically, **weekday_vs_weekend** shows whether the day is a weekday or weekend; **activity_minute** and **activity_number** shows the minute for each activities.

#### Traditional analyses of accelerometer data 
```{r total_activity, warning = F, message = F}
# aggregate activity_minute to get total_activity var. for each day
accelerometer %>% 
    group_by(week,day) %>% 
    summarize(total_activity = sum(activity_minute)) %>% 
    # create a table showing these totals
    pivot_wider(names_from = day,
                values_from = total_activity) %>% 
    knitr::kable()
```

Apparent trend I noticed is that week 4 and 5's Saturday both have the exact same amount of low activity, so I am a little curious whether that two numbers are resulted from reading error or technical error, or the person just happen to move less and has the same amount of activity. Another potential trend is that this person has relatively the same amount of activity on Wednesdays and Thursdays.

#### single-panel plot: 24-hour activity time courses for each day
```{r single-panel plot, warning = F, message = F}
accelerometer %>% 
    ggplot(aes(x = activity_number, 
               y = activity_minute, 
               color = day, group = day_id)) +
    geom_smooth(se = F, method = "loess", geom = "line", alpha = 0.5) +
    labs(title = "24-hour activity time courses for each day",
         x = "Time in a day (in minutes)",
         y = "Activity Count") +
    guides(color = guide_legend(nrow = 1, byrow = T))
```
Based on the above graph, I see the person has more activity during the (middle of the) day, and less/no activity during the early morning and late night. This makes sense as people generally asleep during the early morning and late night, and work during the (middle of the) day. In addition, this person seems to very low (almost 0) activities on one Saturday and Monday, maybe this is due to the person did not (forgot to) put on the accelerometer as him/she started the day.

### Problem 3
First,

```{r load noaa data}
data("ny_noaa") 
```
